---
title: "Project 2"
author: "Coco Donovan"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Necessary Packages

```{r}
library(tidyr)
library(dplyr)
library(readr)
library(reshape2)
library(stringr)
```

## Loading Data Source 1: Student Testing Data (Alice Ding)

```{r}
ds1_link <- 'https://gist.githubusercontent.com/Kimmirikwa/b69d0ea134820ea52f8481991ffae93e/raw/4db7b1698035ee29885d10e1a59bd902716ae168/student_results.csv'

ds1 <- read_csv(ds1_link)
```

## Cleaning Data Source 1

```{r}
ds1 <- ds1 %>%
  melt(id.vars = c('id', 'name', 'phone', 'sex and age','test number')) %>%
  rename(`term number` = variable, score = value)

ds1 <- ds1 %>%
  separate(`sex and age`, into = c("gender", "age"), sep = "_") %>%
  mutate(age = as.numeric(age))

ds1
```

## Converting ds1 to a csv

```{r}
write.csv(ds1, 'Students_Test_Scores.csv')
```

## The suggested analysis for Data Source One

The suggested analysis was, "We could find the average test score for each student or average test score overall. We could also count how many students are of each sex and find the average age as well."

```{r}
# Average Test Score
avg_by_student <- ds1 %>%
  group_by(name) %>%
  summarise(avg_test_score = mean(score)) %>%
  arrange(desc(avg_test_score))

knitr::kable(avg_by_student)

# Counts per gender
gender_counts <- unique(ds1[c("id", "gender")]) %>%
  group_by(gender) %>%
  summarize(count = n())

knitr::kable(gender_counts)

# Averages by gender
avg_by_gender <- ds1 %>%
  group_by(gender) %>%
  summarise(avg_test_score = mean(score)) %>%
  arrange(desc(avg_test_score))

knitr::kable(avg_by_gender)
```

## Loading Data Source 2: Religion and Income

I should note that I did not realize that RStudio has a built in table 
relig_income, but it had more religions than the example I originally provided,
so I just recreated the example I provided for consistency.

```{r}
relig_inc <- read_csv('Raw_Religion_Income.csv')
```

## Cleaning Data Source 2

```{r}
relig_inc <- relig_inc %>%
  melt(id.vars = 'religion') %>%
  mutate(income_level = variable, count = value) %>%
  select(-variable, -value)

relig_inc
```

## Writing relig_inc to a csv

```{r}
write.csv(relig_inc, "Religion_&_Income_Levels.csv")
```

## The suggested analysis for Data Source Two

The suggested analysis was, "As far as analysis goes you could group by religion
and see what the religious makeup of all respondents was by percentages."

```{r}
# Percent of respondents at each income level by religion

percents <- relig_inc %>%
  group_by(religion) %>%
  mutate(percent = round(100 * count/sum(count), 2)) %>%
  ungroup() %>%
  arrange(religion)
```

## Loading Data Source 3: 

```{r}

```



## Conclusion:

